Meta-Analysis of the `Ironic' Effects of Intergroup Contact

### 2021-04-21 manuscript/manuscript.Rmd
update run_scripts.R

---
title: "Meta-Analysis of the 'Ironic' Effects of Intergroup Contact"
bibliography: references.bib
csl: apa.csl
abstract: 
  Critics have argued that intergroup contact, psychology’s most-researched paradigm for reducing prejudice, has the ‘ironic’ effect of stifling support for social change in disadvantaged groups. We conducted a preregistered meta-analytic test of this effect across `r with(dplyr::distinct(readRDS("../data/dl.rds"), sample, id, n), numform::f_comma(dplyr::n_distinct(id)))` studies with `r with(dplyr::distinct(readRDS("../data/dl.rds"), sample, id, n), numform::f_comma(length(sample)))` samples of `r with(dplyr::distinct(readRDS("../data/dl.rds"), sample, id, n), numform::f_comma(sum(n)))` disadvantaged-group members. As predicted, intergroup contact was, on average, associated with less perceived injustice (`r with(readRDS("../results/results_preregistered_analyses.rds"), paste0("$r = ", numform::f_num(median(r_mean[x_var == "ic" & y_var == "pi"]), digits = 2), "$"))`), collective action (`r with(readRDS("../results/results_preregistered_analyses.rds"), paste0("$r = ", numform::f_num(median(r_mean[x_var == "ic" & y_var == "ca"]), digits = 2), "$"))`), and support for reparative policies (`r with(readRDS("../results/results_preregistered_analyses.rds"), paste0("$r = ", numform::f_num(median(r_mean[x_var == "ic" & y_var == "ps"]), digits = 2), "$"))`) in disadvantaged groups. However, these associations were small, variable, and consistent with alterative explanations. Moderator analyses found that these associations were strongest in studies on short-term migration and post-colonial relations, that measured intergroup contact directly, and that involved adults. After controlling for the positive association of negative contact with support for social change, positive contact was not (negatively) associated with any outcome in studies that measured both predictors. We close by highlighting open questions about the relation between intergroup contact and social change.
statement: 
  As societies are becoming more diverse, there are more opportunities for members of different groups to come in contact with each other. Psychologists disagree about what this means for overcoming persistent inequality between advantaged and disadvantaged groups. Some argue that intergroup contact reduces advantaged-group members' prejudice and, thereby, prevents interpersonal discrimination. Reducing prejudice, however, is rarely enough to overcome institutional and structural discrimination. Instead, social change often results from disadvantaged-group members mobilizing against social injustice (for example, in the 2020 Black Lives Matter protests). Critics argue that positive intergroup contact could create a false sense of equality and, thereby, stifle disadvantaged-group members' opposition to injustice. We present the first systematic review and meta-analysis of the growing number of studies examining how intergroup contact relates to support for social change in disadvantaged groups. Our findings help us understand whether, when, and how intergroup contact might hinder, rather than help, social change.
output:
  pdf_document:
    latex_engine: pdflatex
    keep_tex: yes
    template: preprint-template.tex
  word_document:
    reference_docx: word-template.docx
---

@@

We prepared a protocol in accordance with the PRISMA-P statement [@moher_preferred_2015] and preregistered it on the Open Science Framework.^[[https://osf.io/ryuev/?view_only=8938ac6178914da98bdda1da4a6e27f0](https://osf.io/ryuev/?view_only=8938ac6178914da98bdda1da4a6e27f0)] Preregistration precludes undisclosed flexibility in study selection, outcome selection, and other decisions that influence effect size estimates—and thus prevents confirmation bias from determining the conclusions of a meta-analysis [@lakens_reproducibility_2016]. We report deviations from our protocol in the Supplemental Materials. We make data and analysis scripts available online.^[[https://osf.io/w5tqv/?view_only=5cbf7ad5b2b449a7a0f4ada0337b802a](https://osf.io/w5tqv/?view_only=5cbf7ad5b2b449a7a0f4ada0337b802a)]

>>>>

---
title: "Meta-Analysis of the 'Ironic' Effects of Intergroup Contact"
bibliography: references.bib
csl: apa.csl
abstract: 
  Critics have argued that intergroup contact, psychology’s most-researched paradigm for reducing prejudice, has the ‘ironic’ effect of stifling support for social change in disadvantaged groups. We conducted a preregistered meta-analytic test of this effect across `r with(dplyr::distinct(readRDS("../data/dl.rds"), sample, id, n), numform::f_comma(dplyr::n_distinct(id)))` studies with `r with(dplyr::distinct(readRDS("../data/dl.rds"), sample, id, n), numform::f_comma(length(sample)))` samples of `r with(dplyr::distinct(readRDS("../data/dl.rds"), sample, id, n), numform::f_comma(sum(n)))` disadvantaged-group members. As predicted, intergroup contact was, on average, associated with less perceived injustice (`r with(readRDS("../results/results_preregistered_analyses.rds"), paste0("$r = ", numform::f_num(median(r_mean[x_var == "ic" & y_var == "pi"]), digits = 2), "$"))`), collective action (`r with(readRDS("../results/results_preregistered_analyses.rds"), paste0("$r = ", numform::f_num(median(r_mean[x_var == "ic" & y_var == "ca"]), digits = 2), "$"))`), and support for reparative policies (`r with(readRDS("../results/results_preregistered_analyses.rds"), paste0("$r = ", numform::f_num(median(r_mean[x_var == "ic" & y_var == "ps"]), digits = 2), "$"))`) in disadvantaged groups. However, these associations were small, variable, and consistent with alterative explanations. Moderator analyses found that these associations were strongest in studies on short-term migration and post-colonial relations, that measured intergroup contact directly, and that involved adults. After controlling for the positive association of negative contact with support for social change, positive contact was not (negatively) associated with any outcome in studies that measured both predictors. We close by highlighting open questions about the relation between intergroup contact and social change.
statement: 
  As societies are becoming more diverse, there are more opportunities for members of different groups to come in contact with each other. Psychologists disagree about what this means for overcoming persistent inequality between advantaged and disadvantaged groups. Many argue that intergroup contact reduces advantaged-group members' prejudice and, thereby, prevents interpersonal discrimination. Reducing prejudice, however, is rarely enough to overcome institutional and structural discrimination. Instead, social change often results from disadvantaged-group members mobilizing against social injustice (for example, in the 2020 Black Lives Matter protests). Critics argue that positive intergroup contact could create a false sense of equality and, thereby, stifle disadvantaged-group members' opposition to injustice. We present the first systematic review and meta-analysis of the growing number of studies examining how intergroup contact relates to support for social change in disadvantaged groups. Our findings help us understand whether, when, and how intergroup contact might hinder, rather than help, social change.
output:
  pdf_document:
    latex_engine: pdflatex
    keep_tex: yes
    template: preprint-template.tex
  word_document:
    reference_docx: word-template.docx
---

@@

We prepared a protocol in accordance with the PRISMA-P statement [@moher_preferred_2015] and preregistered it on the Open Science Framework.^[[https://osf.io/ryuev/?view_only=8938ac6178914da98bdda1da4a6e27f0](https://osf.io/ryuev/?view_only=8938ac6178914da98bdda1da4a6e27f0)] Preregistration precludes undisclosed flexibility in study selection, outcome selection, and other decisions that influence effect size estimates—and thus prevents confirmation bias from determining the conclusions of a meta-analysis [@lakens_reproducibility_2016]. We report deviations from our protocol in the Supplemental Materials. We make data, analysis scripts, and the fully reproducible manuscript available online.^[[https://osf.io/w5tqv/?view_only=5cbf7ad5b2b449a7a0f4ada0337b802a](https://osf.io/w5tqv/?view_only=5cbf7ad5b2b449a7a0f4ada0337b802a)]

### 2021-04-21 manuscript/manuscript.tex
update run_scripts.R

\twocolumn[
  \begin{@twocolumnfalse}
    \maketitle
    \begin{abstract}
    \noindent Critics have argued that intergroup contact, psychology's
most-researched paradigm for reducing prejudice, has the `ironic' effect
of stifling support for social change in disadvantaged groups. We
conducted a preregistered meta-analytic test of this effect across 96
studies with 137 samples of 211,360 disadvantaged-group members. As
predicted, intergroup contact was, on average, associated with less
perceived injustice (\(r = -.07\)), collective action (\(r = -.06\)),
and support for reparative policies (\(r = -.07\)) in disadvantaged
groups. However, these associations were small, variable, and consistent
with alterative explanations. Moderator analyses found that these
associations were strongest in studies on short-term migration and
post-colonial relations, that measured intergroup contact directly, and
that involved adults. After controlling for the positive association of
negative contact with support for social change, positive contact was
not (negatively) associated with any outcome in studies that measured
both predictors. We close by highlighting open questions about the
relation between intergroup contact and social change.
    \end{abstract}
  \end{@twocolumnfalse}
  \renewcommand{\abstractname}{Statement of Relevance}
  \begin{@twocolumnfalse}
    \maketitle
    \begin{abstract}
    \noindent As societies are becoming more diverse, there are more opportunities for
members of different groups to come in contact with each other.
Psychologists disagree about what this means for overcoming persistent
inequality between advantaged and disadvantaged groups. Some argue that
intergroup contact reduces advantaged-group members' prejudice and,
thereby, prevents interpersonal discrimination. Reducing prejudice,
however, is rarely enough to overcome institutional and structural
discrimination. Instead, social change often results from
disadvantaged-group members mobilizing against social injustice (for
example, in the 2020 Black Lives Matter protests). Critics argue that
positive intergroup contact could create a false sense of equality and,
thereby, stifle disadvantaged-group members' opposition to injustice. We
present the first systematic review and meta-analysis of the growing
number of studies examining how intergroup contact relates to support
for social change in disadvantaged groups. Our findings help us
understand whether, when, and how intergroup contact might hinder,
rather than help, social change.
    \end{abstract}
  \end{@twocolumnfalse}
]

@@

We prepared a protocol in accordance with the PRISMA-P statement (Moher
et al., 2015) and preregistered it on the Open Science
Framework.\footnote{\url{https://osf.io/ryuev/?view_only=8938ac6178914da98bdda1da4a6e27f0}}
Preregistration precludes undisclosed flexibility in study selection,
outcome selection, and other decisions that influence effect size
estimates---and thus prevents confirmation bias from determining the
conclusions of a meta-analysis (Lakens et al., 2016). We report
deviations from our protocol in the Supplemental Materials. We make data
and analysis scripts available online.\footnote{\url{https://osf.io/w5tqv/?view_only=5cbf7ad5b2b449a7a0f4ada0337b802a}}

>>>>

\twocolumn[
  \begin{@twocolumnfalse}
    \maketitle
    \begin{abstract}
    \noindent Critics have argued that intergroup contact, psychology's
most-researched paradigm for reducing prejudice, has the `ironic' effect
of stifling support for social change in disadvantaged groups. We
conducted a preregistered meta-analytic test of this effect across 96
studies with 137 samples of 211,360 disadvantaged-group members. As
predicted, intergroup contact was, on average, associated with less
perceived injustice (\(r = -.07\)), collective action (\(r = -.06\)),
and support for reparative policies (\(r = -.07\)) in disadvantaged
groups. However, these associations were small, variable, and consistent
with alterative explanations. Moderator analyses found that these
associations were strongest in studies on short-term migration and
post-colonial relations, that measured intergroup contact directly, and
that involved adults. After controlling for the positive association of
negative contact with support for social change, positive contact was
not (negatively) associated with any outcome in studies that measured
both predictors. We close by highlighting open questions about the
relation between intergroup contact and social change.
    \end{abstract}
  \end{@twocolumnfalse}
  \renewcommand{\abstractname}{Statement of Relevance}
  \begin{@twocolumnfalse}
    \maketitle
    \begin{abstract}
    \noindent As societies are becoming more diverse, there are more opportunities for
members of different groups to come in contact with each other.
Psychologists disagree about what this means for overcoming persistent
inequality between advantaged and disadvantaged groups. Many argue that
intergroup contact reduces advantaged-group members' prejudice and,
thereby, prevents interpersonal discrimination. Reducing prejudice,
however, is rarely enough to overcome institutional and structural
discrimination. Instead, social change often results from
disadvantaged-group members mobilizing against social injustice (for
example, in the 2020 Black Lives Matter protests). Critics argue that
positive intergroup contact could create a false sense of equality and,
thereby, stifle disadvantaged-group members' opposition to injustice. We
present the first systematic review and meta-analysis of the growing
number of studies examining how intergroup contact relates to support
for social change in disadvantaged groups. Our findings help us
understand whether, when, and how intergroup contact might hinder,
rather than help, social change.
    \end{abstract}
  \end{@twocolumnfalse}
]

@@

We prepared a protocol in accordance with the PRISMA-P statement (Moher
et al., 2015) and preregistered it on the Open Science
Framework.\footnote{\url{https://osf.io/ryuev/?view_only=8938ac6178914da98bdda1da4a6e27f0}}
Preregistration precludes undisclosed flexibility in study selection,
outcome selection, and other decisions that influence effect size
estimates---and thus prevents confirmation bias from determining the
conclusions of a meta-analysis (Lakens et al., 2016). We report
deviations from our protocol in the Supplemental Materials. We make
data, analysis scripts, and the fully reproducible manuscript available
online.\footnote{\url{https://osf.io/w5tqv/?view_only=5cbf7ad5b2b449a7a0f4ada0337b802a}}
